---
output: 
   pdf_document:
header-includes:
   - \usepackage{amssymb, amsmath, amsthm}
   - \usepackage{tabu}
   - \newcommand{\E}{\mathbb{E}}
   - \newcommand{\var}{{\rm Var}}
   - \newcommand{\N}{\mathcal{N}}
---

\noindent \begin{tabu} to \textwidth {@{}X[4 l] @{}X[r]}
  \textbf{Assigment 1}           & \\ 
  \textbf{Mgmt 237E: Empirical Methods} & \\ 
  \textbf{Ian Laker, PrasanthKumar, Nitish Ramkumar}         & 
\end{tabu}

##Problem 1
##1  
  
We know the jump at time t is $J_t = B_t(\mu_j + \sigma_j \delta_t$)  
  
where $\delta_t$ = N(0,1)  
  
and the log returns $r_t = \mu + \sigma \epsilon_t + J_t$  
  
**Mean**  
$\E(J_t) = \E[B_t(\mu_j + \sigma_j \delta_t)]$ = $\E(B_t)\E(\mu_j + \sigma_j \delta_t)$ = $\mathbf{p\mu_j}$   

$\E(r_t) = \E(\mu) + \E(\sigma \epsilon_t) + \E(J_t)$ = $\underline{\mathbf{\mu + p \mu_j}}$
   
**Variance**  
  Var($J_t$) = $\E(J_t^2$) - $(\E(J_t))^2$    
  = $\E({B_t}^2(\mu_j + \sigma_j \delta_t)^2) - {\mu_j}^2 p^2$  
  = $\underline{\E({B_t}^2) \E((\mu_j + \sigma_j \delta_t)^2)} - {\mu_j}^2 p^2$  
    
  $\underline{\E({B_t}^2)} = Var(B_t) + (E(B_t))^2 = p(1-p) + p^2 = p$  
$\underline{\E((\mu_j + \sigma_j \delta_t)^2)} = Var(\mu_j + \sigma_j \delta_t) + (\E(\mu_j + \sigma_j \delta_t))^2 = {{\sigma_j}^2 + {\mu_j}^2}$  
  
  So, Var($J_t$) =  $\mathbf{p {\sigma_j}^2 + p {\mu_j}^2 - {\mu_j}^2 p^2}$  
  
Var($r_t$) = Var($\mu + \sigma \epsilon_t$) + Var($J_t$) = $\underline{\mathbf{{\sigma}^2 + p {\sigma_j}^2 + p {\mu_j}^2 - {\mu_j}^2 p^2}}$  
  
**Skewness**  
  If Skew in terms of mean and variance
  = $\frac{\E({J_t}^3) - 3\E(J_t) Var(J_t) - (\E(J_t))^3}{(Var(J_t))^{3/2}}$   
  Then 3rd moment $\mu_3(J_t)$ = $\E({J_t}^3) - 3\E(J_t) Var(J_t) - (\E(J_t))^3$  
bitch
******  
  $\E({J_t}^3) = \E({B_t}^3(\mu_j + \sigma_j \delta_t)^3) = \underline{\E({B_t}^3)\E((\mu_j + \sigma_j \delta_t)^3)}$  
  
  We can find $\underline{\E({B_t}^3)}$ by using bernoulli distribution mean, variance and skew( $\frac{1-2p}{\sqrt{p(1-p)}}$). Substitute in skew formula in terms of mean and variance   
  
  $\E({B_t}^3) = \frac{(1-2p){(p(1-p))}^{3/2}}{\sqrt{p(1-p)}} + 3p^2(1-p) - p^3$   
= $(1-2p)(p-p^2) + 3p^2 - 3p^3 + p^3 = \mathbf{p}$  
    
  We can find $\underline{\E((\mu_j + \sigma_j \delta_t)^3)}$ by using normal distribution mean, variance and skew (0). Substitute in skew formula in terms of mean and variance  
  
 $\E((\mu_j + \sigma_j \delta_t)^3) = \mathbf{3\mu_j {\sigma_j}^2 + {\mu_j}^3}$  
   
  so $\E({J_t}^3) = \mathbf{p(3\mu_j {\sigma_j}^2 + {\mu_j}^3)}$  
    
so $\mu_3(J_t) = \mathbf{3\mu_j{\sigma_j}^2 p + {\mu_j}^3 p - 3 p^2 \mu_j {\sigma_j}^2 - 3 p^2 {\mu_j}^3 + 2 p^3 {\mu_j}^3}$  
    
******          
Skew($r_t$) = Skew($\mu + \sigma \epsilon_t + J_t$)

We know, Skew(A + B) = $\frac{\mu_3(A) + \mu_3(B)}{(Var(A) + Var(B))^{3/2}}$,  
  
Substitute A = $\mu + \sigma \epsilon_t$ and B = $J_t$  
  
Skew($r_t$) = $\frac{0 + \mu_3(B)}{{\sigma}^2 + p {\sigma_j}^2 + p {\mu_j}^2 - {\mu_j}^2 p^2)^{3/2}}$  
  
  = $\underline{\mathbf{\frac{3\mu_j{\sigma_j}^2 p + {\mu_j}^3 p - 3 p^2 \mu_j {\sigma_j}^2 - 3 p^2 {\mu_j}^3 + 2 p^3 {\mu_j}^3}{({\sigma}^2 + p {\sigma_j}^2 + p {\mu_j}^2 - {\mu_j}^2 p^2)^{3/2}}}}$

  
##2 
The general stock returns are distributed with skew and thick tails. A log normal distribution doesn't cover the skew and the thick tails which are necessary to represent general stock returns.

##3

Get the SP500 monthly data to use as baseline
```{r}
# Function to get data #
getData <- function(sql, n = -1){
  #setup connection
  res <- dbSendQuery(wrds, sql)
  dbHasCompleted(res)
  
  #perform fetch
  returnData <- fetch(res, n)
  
  #clear memory
  dbClearResult(res)
  return(returnData)
}


sql3 <- "SELECT * FROM CRSP.MSP500"
msp500.all <- getData(sql3)
msp500.all$caldt <- as.Date(msp500.all$caldt)
msp500.all.xts <- xts::xts(log(1+msp500.all$vwretd[-1]),order.by = msp500.all$caldt[-1])
colnames(msp500.all.xts) <- "vwretd"
```

If the returns are normally distributed, the distribution will look like this compared to the monthly SP returns.
```{r}
plot(density(msp500.all.xts),col="blue", type="l",xlim=c(-0.2,0.2))

#only normal
lines(density(rnorm(n = 600,mean = 0.008,sd = 0.063)),col="red",type="l")
```

If we take the bernoulli-normal mix, the distribution will look like this compared to the monthly SP returns. 

```{r}
#normalBernoullimix
normalBernoulliMix <- function(normalMean,normalSD,bernProb,jumpMean,jumpSD,n)
{
  SecondTerm <- jumpMean + jumpSD*rnorm(n)
  jt <- rbinom(n,1,bernProb)*(SecondTerm)
  normalMean + normalSD * rnorm(n) + jt
}

normalBernoulliDist <- normalBernoulliMix(0.012,0.05,0.15,-0.03,0.1,600)  

#mean, SD, skewness, Kurtosis
mean(normalBernoulliDist)
sd(normalBernoulliDist)
library(moments)
skewness(normalBernoulliDist)
kurtosis(normalBernoulliDist)

#plot sp500
plot(density(msp500.all.xts),col="blue", type="l")

#density
lines(density(normalBernoulliDist),type="l",col="red", ylim=c(0,9))
#lines(density(rnorm(600,0.012,0.05)),col="red")

#actual values
lines(normalBernoulliDist,type="l",col="blue")
#lines(rnorm(600,0.012,0.05),type="l",col="red")

```

##Problem 2

Get the necessary data
```{r}
#this.dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
setwd("C:/_UCLA/237E_Empirical/Assigments/Assignment1")
library("readxl")
library("xts")
head(dbv) <- read_excel("DBV.xlsx")
gspc <- read_excel("GSPC.xlsx")

dbv$Date <- as.Date(dbv$Date)
gspc$Date <- as.Date(gspc$Date)
dbv.xts <- xts(dbv[,-1],order.by = dbv$Date)
gspc.xts <- xts(gspc[,-1],order.by = gspc$Date)
```

Calculate the log returns
```{r}
dbv.logreturns <- log(dbv.xts$'Adj Close'[-1,]/lag(dbv.xts$'Adj Close')[-1,])
gspc.logreturns <- log(gspc.xts$'Adj Close'[-1,]/lag(gspc.xts$'Adj Close')[-1,])
```

##1
```{r}
#1
plot(dbv.logreturns)
plot(gspc.logreturns)

hist(dbv.logreturns)
hist(gspc.logreturns)
```

##2
##a
```{r}
skewNullCheck <- function(returns,alpha=.05){
  criticalt <- qt(1-(alpha/2),df = length(returns))
  skewcap <- skewness(returns)
  skewt <- skewcap/(sqrt(6/length(returns)))
  returnVal <- c(skewcap,skewt,abs(skewt) > criticalt) #TRUE, so reject normal distribution, no skewness Null
  names(returnVal) <- c("Sample Skewness","Skewness t","Reject Null?")
  returnVal
}

skewNullCheck(dbv.logreturns)
skewNullCheck(gspc.logreturns)

#b
kurtosisNullCheck <- function(returns,alpha=.05){
  criticalt <- qt(1-(alpha/2),df = length(returns))
  kurtosiscap <- kurtosis(returns)
  kurtosist <- (kurtosiscap-3)/(sqrt(24/length(returns)))
  returnVal <- c(kurtosiscap,kurtosist,abs(kurtosist) > criticalt) #TRUE, so reject normal distribution, no skewness Null
  names(returnVal) <- c("Sample Kurtosis","Kurtosis t","Reject Null?")
  returnVal
}

kurtosisNullCheck(dbv.logreturns)
kurtosisNullCheck(gspc.logreturns)

#c
jbTest <- function(returns,alpha=.05){
  criticalchi <- qchisq(1-alpha,df = 2)
  skewcap <- skewness(returns)
  kurtosiscap <- kurtosis(returns)
  jbt <- skewcap^2/(6/length(returns)) + (kurtosiscap-3)^2/(24/length(returns))
  abs(jbt) > criticalchi #TRUE, so reject normal distribution, no skewness Null
}

jbTest(dbv.logreturns)
jbTest(gspc.logreturns)
```


##3
```{r}
skewKurtMat <- matrix(c(skewNullCheck(dbv.logreturns)[1],skewNullCheck(gspc.logreturns)[1],kurtosisNullCheck(dbv.logreturns)[1],kurtosisNullCheck(gspc.logreturns)[1]),nrow = 2)
colnames(skewKurtMat) <- c("Skewness","Kurtosis")
row.names(skewKurtMat) <- c("DBV","GSPC")
skewKurtMat
```

##4
The expected return to standard deviation ratio covers only the first 2 moments of the return. It doesn't show cover the difference in skewness and kurtosis between the two investments.  
As GSPC has lower negative skewness, there is lesser chance of getting a negative tail value.  
It also has lower kurtosis, which means there is lesser chance of getting an extreme value  

##5
##a 
Both assumptions valid (i.e. homoskedastic and normal)

```{r}
reg1Summ <- summary(lm(dbv.logreturns ~ gspc.logreturns))
reg1Summ$coefficients[,2]

library(DataAnalytics)
reg2Summ <- lmSumm(lm(dbv.logreturns ~ gspc.logreturns),HAC = T)
reg2Summ$coef.table[,2]


```



