---
output: 
   pdf_document:
header-includes:
   - \usepackage{amssymb, amsmath, amsthm}
   - \usepackage{tabu}
   - \usepackage{physics}
   - \newcommand{\E}{\mathbb{E}}
   - \newcommand{\N}{\mathcal{N}}
---

\noindent \begin{tabu} to \textwidth {@{}X[4 l] @{}X[r]}
  \textbf{Assigment 1}           & \\ 
  \textbf{Mgmt 237E: Empirical Methods} & \\ 
  \textbf{Ian Laker, PrasanthKumar, Nitish Ramkumar}         & 
\end{tabu}

##Problem 1
##1  

Let us generate the moment generating function for $r_t$  
  
Moment generating function Mg(s) = $\E(exp{r_{t}s})$  
  
= $\E(e^{[\mu + \sigma \epsilon_t + \beta_t(\mu_j + \sigma_j \delta_t)]s})$  
  
= $\E(e^{(\mu + \sigma \epsilon_t)s} e^{(\beta_t(\mu_j + \sigma_j \delta_t))s})$  
  
= $\E(e^{(\mu + \sigma \epsilon_t)s}) \E(e^{\beta_t(\mu_j + \sigma_j \delta_t)s})$  
  
= $e^{s \mu + \frac{1}{2} \sigma^2 s^2} (p  \E(e^{(\mu_j + \sigma_j \delta_t)s}) + (1-p)e^{0})$  
  
= $\mathbf{e^{s \mu + \frac{1}{2} \sigma^2 s^2} [p  e^{s \mu_j + \frac{1}{2} {\sigma_j}^2 s^2} + (1-p)]}$  

**Mean**

Lets keep A = $s \mu + \frac{1}{2} \sigma^{2} s^2 + s \mu_j + \frac{1}{2} {\sigma_j}^2 s^2$, B = $s \mu + \frac{1}{2} \sigma^2 s^2$,  
  
so Mg(s) = $p e^A + (1-p) e^B$  
  
We know Mean = 1st Moment ($M_1$) = $\frac{\partial Mg}{\partial s}(s=0)$  
  
$\frac{\partial Mg}{\partial s} = p e^A \frac{\partial A}{\partial s} + (1-p) e^B \frac{\partial B}{\partial s}$  
  
We can calculate $\frac{\partial A}{\partial s}(s=0) = \mu + \mu_j,  \frac{\partial B}{\partial s}(s=0) = \mu$  
  
$Mean (\mu_{r_t}) = M_1 = \frac{\partial Mg}{\partial s}(s=0) = \mathbf{ \mu + p \mu_j}$
  
    
**Variance**

We know Variance = $M_2$ - ${\mu_{r_t}}^2$

$M_2 = \frac{\partial^2 Mg}{\partial s^2}(s=0)$  
  
$\frac{\partial^2 Mg}{\partial s^2} = pe^A{\frac{\partial A}{\partial s}}^2 + pe^A\frac{\partial^2 A}{\partial s^2} + e^B{\frac{\partial B}{\partial s}}^2 + e^B\frac{\partial^2 B}{\partial s^2} - pe^A{\frac{\partial B}{\partial s}}^2 - pe^B\frac{\partial^2 B}{\partial s^2}$  
  
We can calculate $\frac{\partial^2 A}{\partial s^2}(s=0) = \sigma^2 + {\sigma_j}^2,  \frac{\partial^2 B}{\partial s^2}(s=0) = \sigma^2$

So, $M_2 = p{\mu_j}^2 + 2p\mu \mu_j + p \sigma^2 + p {\sigma_j}^2 + \mu^2 + \sigma^2$  
  
Variance $({\sigma_{r_t}}^2) = \mathbf{\sigma^2 + p {\sigma_j}^2 + p {\mu_j}^2 - p^2 {\mu_j}^2}$  
  
      
**Skewness**  
  
We know Skewness = $\frac{M_3 - 3 \mu_{r_t} {\sigma_{r_t}}^2 - {\mu_{r_t}}^3}{{\sigma_{r_t}}^{\frac{3}{2}}}$  
  
$M_3 = \pdv[3]{Mg}{s}(s=0)$  
  
$\frac{\partial^3 Mg}{\partial s^3} = pe^A {\frac{\partial A}{\partial s}}^3 + 3pe^A \frac{\partial A}{\partial s} \frac{\partial^2 A}{\partial s^2}  + p e^A \frac{\partial^3 A}{\partial s^3} + e^B {\frac{\partial B}{\partial s}}^3 + 3e^B \frac{\partial B}{\partial s} \frac{\partial^2 B}{\partial s^2}  +  e^B \frac{\partial^3 B}{\partial s^3} - pe^B (\frac{\partial B}{\partial s})^3 - 3pe^B \frac{\partial B}{\partial s} \frac{\partial^2 B}{\partial s^2}  - p e^B \frac{\partial^3 B}{\partial s^3}$  
    
We can calculate $\frac{\partial^3 A}{\partial s^3}(s=0) = 0,  \frac{\partial^3 B}{\partial s^3}(s=0) = 0$  
  
$M_3 = p(\mu + \mu_j)^3 + 3p(\mu + \mu_j)(\sigma^2 + {\sigma_j}^2) + (1-p)(\mu^3 + 3 \mu \sigma^2)$  
  
So Skewness = $\frac{p{\mu_j}^3 + 3p\mu_j{\sigma_j}^2 - 3p^2\mu_j {\sigma_j}^2 - 3 p^2 {\mu_j}^3 + 2p^3{\mu_j}^3}{(\sigma^2 + p {\sigma_j}^2 + p {\mu_j}^2 - p^2 {\mu_j}^2)^{\frac{3}{2}}}$  
  
= $\mathbf{\frac{\mu_j p({\mu_j}^2(2p^2 -3p + 1) - 3 {\sigma_j}^2(p-1))}{(\sigma^2 + p {\sigma_j}^2 + p {\mu_j}^2 - p^2 {\mu_j}^2)^{\frac{3}{2}}}}$  
  
**Kurtosis**  
  
We know Kurtosis = $\frac{M_4 - 4 \mu_{r_t}M_3 + 6 {\mu_{r_t}}^2 M_2 - 3 {\mu_{r_t}}^4}{{\sigma_{r_t}}^2}$  
  
$M_4 = \pdv[4]{Mg}{s}(s=0)$  
  
$\pdv[4]{Mg}{s} = p e^A {\pdv[1]{Mg}{s}}^4 + 6 p e^A {\pdv[1]{Mg}{s}}^2 \pdv[2]{Mg}{s} + 4p e^A \pdv[1]{Mg}{s} \pdv[3]{Mg}{s} + 3p e^A {\pdv[2]{Mg}{s}}^2 + p e^A \pdv[4]{Mg}{s}$  
  
$\pdv[4]{A}{s} = 0, \pdv[4]{B}{s} = 0$  
  
$M_4 = p(\mu + \mu_j)^4 + 6p(\mu + \mu_j)^2(\sigma^2 + {\sigma_j}^2) + 3p(\sigma^2 + {\sigma_j}^2)^2 + (1-p)(\mu^4 + 6\mu^2 \sigma^2 + 3 \sigma^3)$  
  
So Kurtosis = $\mathbf{\frac{3 {\sigma_j}^4 p (1 - p) + {\mu_j}^4 p (-6p^3 + 12p^2 -7p +1) + {\mu_j}^3 {\sigma_j}^3 p (12p^2 - 18p _ 6)}{(\sigma^2 + p {\sigma_j}^2 + p {\mu_j}^2 - p^2 {\mu_j}^2)^2}}$ 
  
##2 
The general stock returns are distributed with skew and thick tails. A log normal distribution doesn't cover the skew and the thick tails which are necessary to represent general stock returns.

##3

Get the SP500 monthly data to use as baseline
```{r}
# Function to get data #
getData <- function(sql, n = -1){
  #setup connection
  res <- dbSendQuery(wrds, sql)
  dbHasCompleted(res)
  
  #perform fetch
  returnData <- fetch(res, n)
  
  #clear memory
  dbClearResult(res)
  return(returnData)
}


sql3 <- "SELECT * FROM CRSP.MSP500"
msp500.all <- getData(sql3)
msp500.all$caldt <- as.Date(msp500.all$caldt)
msp500.all.xts <- xts::xts(log(1+msp500.all$vwretd[-1]),order.by = msp500.all$caldt[-1])
colnames(msp500.all.xts) <- "vwretd"

```

If the returns are normally distributed, the distribution will look like this compared to the monthly SP returns.
```{r}
plot(density(msp500.all.xts),col="blue", type="l",xlim=c(-0.3,0.3))

#only normal
lines(density(rnorm(n = 600,mean = 0.008,sd = 0.063)),col="red",type="l")
```

If we take the bernoulli-normal mix, the distribution will look like this compared to the monthly SP returns. 

```{r}
#normalBernoullimix
normalBernoulliMix <- function(normalMean,normalSD,bernProb,jumpMean,jumpSD,n)
{
  SecondTerm <- jumpMean + jumpSD*rnorm(n)
  jt <- rbinom(n,1,bernProb)*(SecondTerm)
  normalMean + normalSD * rnorm(n) + jt
}

normalBernoulliDist <- normalBernoulliMix(0.012,0.05,0.15,-0.03,0.1,600)  

#mean, SD, skewness, Kurtosis
mean(normalBernoulliDist)
sd(normalBernoulliDist)
library(moments)
skewness(normalBernoulliDist)
kurtosis(normalBernoulliDist)

#plot sp500
plot(density(msp500.all.xts),col="blue", type="l")

#density
lines(density(normalBernoulliDist),type="l",col="red", ylim=c(0,9))

#actual values
lines(normalBernoulliDist,type="l",col="blue")

```

##Problem 2

Get the necessary data
```{r}
#this.dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
setwd("C:/_UCLA/237E_Empirical/Assigments/Assignment1")
library("readxl")
library("xts")
dbv <- read_excel("DBV.xlsx")
gspc <- read_excel("GSPC.xlsx")

dbv$Date <- as.Date(dbv$Date)
gspc$Date <- as.Date(gspc$Date)
dbv.xts <- xts(dbv[,-1],order.by = dbv$Date)
gspc.xts <- xts(gspc[,-1],order.by = gspc$Date)
```

Calculate the log returns
```{r}
dbv.logreturns <- log(dbv.xts$'Adj Close'[-1,]/lag(dbv.xts$'Adj Close')[-1,])
gspc.logreturns <- log(gspc.xts$'Adj Close'[-1,]/lag(gspc.xts$'Adj Close')[-1,])
```

##1
```{r}
#1
plot(dbv.logreturns)
plot(gspc.logreturns)

hist(dbv.logreturns,breaks = 50)
hist(gspc.logreturns, breaks = 50)

```

##2
##a
```{r}
skewNullCheck <- function(returns,alpha=.05){
  criticalt <- qt(1-(alpha/2),df = length(returns))
  skewcap <- skewness(returns)
  skewt <- skewcap/(sqrt(6/length(returns)))
  returnVal <- c(skewcap-3,skewt,abs(skewt) > criticalt) #TRUE, so reject normal distribution, no skewness Null
  names(returnVal) <- c("Sample Skewness","Skewness t","Reject Null?")
  returnVal
}

skewNullCheck(dbv.logreturns)
skewNullCheck(gspc.logreturns)

#b
kurtosisNullCheck <- function(returns,alpha=.05){
  criticalt <- qt(1-(alpha/2),df = length(returns))
  kurtosiscap <- kurtosis(returns)
  kurtosist <- (kurtosiscap-3)/(sqrt(24/length(returns)))
  returnVal <- c(kurtosiscap-3,kurtosist,abs(kurtosist) > criticalt) #TRUE, so reject normal distribution, no skewness Null
  names(returnVal) <- c("Sample Kurtosis","Kurtosis t","Reject Null?")
  returnVal
}

kurtosisNullCheck(dbv.logreturns)
kurtosisNullCheck(gspc.logreturns)

#c
jbTest <- function(returns,alpha=.05){
  criticalchi <- qchisq(1-alpha,df = 2)
  skewcap <- skewness(returns)
  kurtosiscap <- kurtosis(returns)
  jbt <- skewcap^2/(6/length(returns)) + (kurtosiscap-3)^2/(24/length(returns))
  returnVal <- c(jbt,abs(jbt) > criticalchi)
  names(returnVal) <- c("JBt","Reject Null?")
  returnVal
}

jbTest(dbv.logreturns)
jbTest(gspc.logreturns)

```


##3
```{r}
sqlDaily <- "SELECT caldt, vwretd FROM CRSPQ.DSP500 WHERE year(caldt) >= 1973 AND year(caldt) < 2015"
dsp500 <- getData(sqlDaily)

skewKurtMat <- matrix(c(skewNullCheck(dbv.logreturns)[1],skewNullCheck(gspc.logreturns)[1],kurtosisNullCheck(dbv.logreturns)[1],kurtosisNullCheck(gspc.logreturns)[1]),nrow = 2)
colnames(skewKurtMat) <- c("Skewness","Kurtosis")
row.names(skewKurtMat) <- c("DBV","GSPC")
skewKurtMat
```

##4
The expected return to standard deviation ratio covers only the first 2 moments of the return. It doesn't show the difference in skewness and kurtosis between the two investments.  
As GSPC has lower negative skewness, there is lesser chance of getting a negative tail value.  
It also has lower kurtosis, which means there is lesser chance of getting an extreme value  

##5
##a 
Both assumptions valid (i.e. homoskedastic and normal)

```{r}
lm <- lm(dbv.logreturns ~ gspc.logreturns)
reg1Summ <- summary(lm)
reg1Summ$coefficients[,2]
vcov(lm)

library(DataAnalytics)
reg2Summ <- lmSumm(lm(dbv.logreturns ~ gspc.logreturns),HAC = T)
reg2Summ$coef.table[,2]


```



