---
title: "Empirical_Assignment5"
output: pdf_document
---

##1
```{r}
#Sample Mean
Avg.Mkt <- 0.05

Avg.1 <- 0.01 + 0.9*Avg.Mkt
Avg.2 <- -0.015 + 1.2*Avg.Mkt
Avg.3 <- 0.005 + 1.0*Avg.Mkt
Avg <- c(Avg.1,Avg.2,Avg.3)
Avg
```

```{r}
#Standard Deviation
Variance_matrix <- matrix(c(0.1^2,0,0,0,0.15^2,0,0,0,0.05^2),nrow=3)
Variance.Mkt <- 0.15^2
Variance.1 <- Variance.Mkt +  Variance_matrix[1,1] 
Variance.2 <- Variance.Mkt +  Variance_matrix[2,2]
Variance.3 <- Variance.Mkt +  Variance_matrix[3,3]
Sd <- c(sqrt(Variance.1),sqrt(Variance.2),sqrt(Variance.3))
Sd
```

```{r}
#SharpeRatio
Avg/Sd
```


##2
To hedge out the market risk, we go short the market based on the beta of the market provided in the regression equation.  
  
```{r}
#Sample Mean
Avg.Hedged.1 <- 0.01
Avg.Hedged.2 <- -0.015
Avg.Hedged.3 <- 0.005
Avg.Hedged <- c(Avg.Hedged.1,Avg.Hedged.2,Avg.Hedged.3)
Avg.Hedged
```

```{r}
#Standard Deviation
Variance.Hedged.1 <- Variance_matrix[1,1] 
Variance.Hedged.2 <- Variance_matrix[2,2]
Variance.Hedged.3 <- Variance_matrix[3,3]
Sd.Hedged <- c(sqrt(Variance.Hedged.1),sqrt(Variance.Hedged.2),sqrt(Variance.Hedged.3))
Sd.Hedged
```

```{r}
#SharpeRatio
Avg.Hedged/Sd.Hedged
```

##3

The maximum sharpe ratio squared based on the mean variance efficiency = $(\bar{R}^e)' \Omega^{-1} \bar{R}^e$  
  
*Proof*  

The aim is to minimize portfolio variance ($w' \Omega w$, where w is the weights and $\Omega$ is variance-covariance matrix), such that the portfolio returns reach the necessary value of *m*  
  
The objective function from the lagrangian form is   
  
*min* $\frac{1}{2} w' \Omega w - k (w' \bar{R}^e - m)$  
  
First order differential w.r.t w  and set it to 0 to minimize
  
$\Omega w - k \bar{R}^e$ = 0  
  
so $w^{MVE} = k \Omega^{-1} \bar{R}^e$  
  
so, ${\bar{R}^e}_{MVE} = (w^{MVE})' \bar{R}^e = k (\bar{R}^e)' \Omega^{-1} \bar{R}^e$
    
$var({R^e}_{MVE}) = (w^{MVE})' \Omega w^{MVE} = k^2 (\bar{R}^e)' \Omega^{-1} \Omega \Omega^{-1} \bar{R}^e$  
= $k^2 (\bar{R}^e)' \Omega^{-1} \bar{R}^e$
  
  
So, the Sharpe Ratio squared for MVE is  
${SR^2}_{MVE}  =  \frac{{\bar{R}^e}_{MVE}}{var({R^e}_{MVE})} = (\bar{R}^e)' \Omega^{-1} \bar{R}^e$  

```{r}
SharpeRatioSq.Max <- t(Avg.Hedged)%*%chol2inv(chol(Variance_matrix))%*%Avg.Hedged
sqrt(SharpeRatioSq.Max)
```

##4
```{r}
SharpeRatio.Market <- 1/3
SharpeRatio.Combined <- sqrt(SharpeRatioSq.Max + SharpeRatio.Market^2)
SharpeRatio.Combined
```

##5a
```{r}
Avg.Hedged.Market <- c(Avg.Mkt,Avg.Hedged)
Variance_matrix_combined <- rbind(c(Variance.Mkt,0,0,0),cbind(rep(0,3),Variance_matrix))

SharpeRatio.Max.Combined <- t(Avg.Hedged.Market)%*%chol2inv(chol(Variance_matrix_combined))%*%Avg.Hedged.Market
Sd <- 0.15
k <- Sd/sqrt(SharpeRatio.Max.Combined)

weights_combined <- (chol2inv(chol(Variance_matrix_combined))%*%Avg.Hedged.Market) * as.numeric(k)
weights_combined
```

##5b
```{r}
#Mean
Mean5 <- Avg.Hedged.Market%*%weights_combined

#SD
Sd5 <- sqrt(t(weights_combined)%*%Variance_matrix_combined%*%weights_combined)

#Sharpe Ratio
Mean5/Sd5

```

##6
#a
```{r}
betas <- c(0.9,1.2,1)
mimick.Return <- ((betas - mean(betas))/(length(betas)*var(betas)))%*%Avg
mimick.Sd <- Variance.Mkt^0.5
mimick.sharpe <- mimick.Return/mimick.Sd
c(mimick.Return,mimick.Sd,mimick.sharpe)
```

#b
```{r}
corr.mimick.market <- mimick.Return * var(betas)
corr.mimick.market
```

#c
```{r}
var.matrix <- Variance_matrix 
var.matrix[1,1] = var.matrix[1,1] + Variance.Mkt
var.matrix[2,2] = var.matrix[2,2] + Variance.Mkt
var.matrix[3,3] = var.matrix[3,3] + Variance.Mkt

eigens <- eigen(cov(var.matrix))
eigens$values/sum(eigens$values)

```

#d
```{r}
eigens$vectors
```

#e
The first PCA is the market. As it can be noticed, though the first PCA explains 66.7\%, the second and third PCA still explains a significant portion of the variance. Also the second PCA is close to a long short strategy, which means that without any exposure to market, it is explaining variance of the returns.  
  