---
title: "Empirical Assignment 6"
author: "Nitish Ramkumar, Ian Laker, Prasanth Kumar"
output: pdf_document
---

#Portfolios Sort
```{r echo=FALSE}
suppressMessages(library(xts))

stockreturns <- matrix(c(2012,-0.12,0.29,0.15,-0.03,2013,-0.03,0.15,0.03,0.05,2014,0.07,0.05,-0.1,0.25,2015,-0.06,-0.01,0.03,0.1,2016,0.33,-0.29,0.08,-0.18),nrow=5,byrow = TRUE)
stockreturns.xts <- as.xts(stockreturns[,-1],as.Date(paste0(stockreturns[,1],"-01-01")))
colnames(stockreturns.xts) <- c("Stock A","Stock B","Stock C","Stock D")

findMedian <- function(vec){
  vec.sorted <- sort(vec)
  if(length(vec.sorted) %% 2 == 0){
    (vec.sorted[length(vec.sorted)/2] + vec.sorted[(length(vec.sorted)/2)+1])/2
  }
  else{
    vec.sorted[length(vec.sorted)%/%2 + 1]
  }
}

```


##1a

The winner stocks based on the sorting strategy are as follows  
```{r echo=FALSE}
#Store positions of winning stocks and convert it into xts
Winner.positions <- xts(t(apply(stockreturns.xts,1,function(x){which(x > findMedian(x))})),order.by=index(stockreturns.xts))

#PRINT WINNERS
Winner.names <- t(sapply(1:nrow(stockreturns.xts),
                        function(count){
                        #Print column name of calculated indices
                        colnames(stockreturns.xts[count,Winner.positions[count]])
                }))
Winner.names <- xts(Winner.names,index(stockreturns.xts))
Winner.names
```
  
The time series returns for the winners and losers are as below
```{r echo=FALSE}
#CALCULATE RETURNS
WinnerLoserReturnsFunc <- function(row){
                          #get previous winners using lag
                          previousWinners <- stats::lag(Winner.positions)[-1][index(row),]
                          #calculate mean of winners(containing previousWinners) and losers
                          c(mean(row[,previousWinners]),mean(row[,-previousWinners]))
                        }

WinnerLoserReturns <- t(sapply(1:nrow(stockreturns.xts[-1]),function(count)
                                    {
                                      WinnerLoserReturnsFunc(stockreturns.xts[-1][count,])
                                    }))

colnames(WinnerLoserReturns) <- c("Winner","Loser")
WinnerLoserReturns <- xts(WinnerLoserReturns,index(stockreturns.xts[-1]))
WinnerLoserReturns
```
  
The mean returns for the winners and losers are as below 
```{r echo=FALSE}
#MEAN OF RETURNS
apply(WinnerLoserReturns,2,mean)
```

##1b
Yes, we can see the expected momentum result, as the winners gain substantially more than the losers.
  
The average market returns is *3.8\%*   
```{r echo=FALSE}
Market.weights <- c(0.3,0.2,0.353,0.147)
Market.returns <- stockreturns.xts%*%Market.weights

#MEAN MARKET 
Market.Mean <- mean(Market.returns)
```

##1c  
The time series returns after creating the long short portfolio is
```{r echo=FALSE}
LongShort.Returns <- WinnerLoserReturns[,'Winner'] - WinnerLoserReturns[,'Loser']
colnames(LongShort.Returns) <- "LongShort"
LongShort.Returns
```

```{r echo=FALSE}
#MEAN
LongShort.mean <- mean(LongShort.Returns)
```
The average long-short returns is **4.625\%**  

```{r echo=FALSE}
#REGRESSION
reg <- summary(lm(LongShort.Returns ~ Market.returns[-1]))
betaEst <- reg$coefficients[2,1]
```
When we run a regression between the long short returns and the market returns, we get a correlation (market beta) of **0.003169**. Given this small value of beta, we can say that CAPM doesn't hold good, as explained returns are not correlated to the market.  
  

##2

The time series returns from the quartile sorts is

```{r echo=FALSE}
Positions<- t(apply(stockreturns.xts,1,
                        function(timeVal){
                            sortedVal <- sort(coredata(timeVal))
                            sapply(sortedVal,function(elem){which(timeVal==elem)})
                          }))
Positions.xts <- xts(Positions,index(stockreturns.xts))
Positions.xts.lag <- stats::lag(Positions.xts)[-1]

QuartileSorts <- t(sapply(1:nrow(Positions.xts.lag),function(x){
  val <- stockreturns.xts[-1][x,Positions.xts.lag[x,]]
  colnames(val) <- NULL
  val
  }))

colnames(QuartileSorts) <- c("Sort1","Sort2","Sort3","Sort4")
QuartileSorts <- xts(QuartileSorts,index(Positions.xts.lag))
QuartileSorts
```

The mean of the quartile sorts is
```{r echo=FALSE}
#MEAN
apply(QuartileSorts,2,mean)
```

The expected momentum result is not achieved as the maximum quartile in this momentum strategy has return less than the minimum quartile.

#AR(p) Process  

##3a 
  
The plot for the AR(2) plot, given $\phi_1$ and $\phi_2$ is    
```{r echo=FALSE}
phi1 <- 1.1
phi2 <- -0.25

set.seed('1234')
ar.sim1 <-arima.sim(model=list(ar=c(phi1,phi2)),n=1000) 
acf(ar.sim1,lag.max = 20,main="Autocorrelation plot for AR(2)")
```

##3b
```{r echo=FALSE}
x1 = (phi1 + sqrt(phi1^2 + 4*phi2))/(-2 * phi2)
x2 = (phi1 - sqrt(phi1^2 + 4*phi2))/(-2 * phi2)

root1 <- 1/x1
root2 <- 1/x2

```
As modulus of both characteristic roots of $1 - \phi_1 x - \phi_2 x^2 = 0$ (0.32087,0.77913) are less than 1, this is a **stationary process**  
  
##3c
Let's assume $r_{t+6} - \mu = x_{t+6}$. As $\mu$ is mean for all elements of the time series, this can be extended to $x_{t+5}, x_{t+4}$ as well.  

To get the dynamic multiplier, we can start with $x_{t+6} = \phi_1 x_{t+5} + \phi_2 x_{t+4} + \epsilon_{t+6}$, and then substitute $x_{t+5}$ and $x_{t+4}$ with further lag values.  
  
$= \epsilon_{t+6} + \phi_1 \epsilon_{t+5} + ({\phi_1}^2 + \phi_2) x_{t+4} + \phi_1 \phi_2 x_{t+3}$  
  
$= \epsilon_{t+6} + \phi_1 \epsilon_{t+5} + ({\phi_1}^2 + \phi_2) \epsilon_{t+4} + ({\phi_1}^3 + 2 \phi_1 \phi_2) x_{t+3} + ({\phi_1}^2 \phi_2 + {\phi_2}^2) x_{t+2}$     
  
$= \epsilon_{t+6} + \phi_1 \epsilon_{t+5} + ({\phi_1}^2 + \phi_2) \epsilon_{t+4} + ({\phi_1}^3 + 2 \phi_1 \phi_2) \epsilon_{t+3} + ({\phi_1}^4 + 3{\phi_1}^2\phi_2 + {\phi_2}^2) x_{t+2} + ({\phi_1}^3 \phi_2 + 2 \phi_1 {\phi_2}^2) x_{t+1}$  
  
$= ....({\phi_1}^4 + 3{\phi_1}^2\phi_2 + {\phi_2}^2) \epsilon_{t+2} + ({\phi_1}^5 + 4 {\phi_1}^3 \phi_2 + 3 \phi_1 {\phi_2}^2) x_{t+1} + ({\phi_1}^4 \phi_2 + 3 {\phi_1}^2 {\phi_2}^2 + {\phi_2}^3) x_t$  
  
$= .....({\phi_1}^5 + 4 {\phi_1}^3 \phi_2 + 3 \phi_1 {\phi_2}^2) \epsilon_{t+1} + ({\phi_1}^6 + 5 {\phi_1}^4 \phi_2 + 6 {\phi_1}^2 {\phi_2}^2 + {\phi_2}^3) x_t + ({\phi_1}^5 \phi_2 + 4 {\phi_1}^3 {\phi_2}^3 + 3 \phi_1 {\phi_2}^3) x_{t-1}$  
  
$=.....({\phi_1}^6 + 5 {\phi_1}^4 \phi_2 + 6 {\phi_1}^2 {\phi_2}^2 + {\phi_2}^3) \epsilon_t + ...$  
  
So, Dynamic multiplier for 6 period ago shock = $\frac{\partial x_{t+6}}{\partial \epsilon_t}$ =   
  
  $=\mathbf{{\phi_1}^6 + 5 {\phi_1}^4 \phi_2 + 6 {\phi_1}^2 {\phi_2}^2 + {\phi_2}^3}$

```{r echo=FALSE}
dynamicMult_1 <- phi1^6 + 5*phi1^4*phi2 + 6*phi1^2*phi2^2 + phi2^3
```

We know, $\phi_1$ = 1.1, $\phi_2$ = -0.25,  So Dynamic Multiplier is **0.379561**  
  
##3d
```{r echo=FALSE}
phi1 = 0.9
phi2 = 0.8

dynamicMult_2 <- phi1^6 + 5*phi1^4*phi2 + 6*phi1^2*phi2^2 + phi2^3

x1_2 = (phi1 + sqrt(phi1^2 + 4*phi2))/(-2 * phi2)
x2_2 = (phi1 - sqrt(phi1^2 + 4*phi2))/(-2 * phi2)

root1_2 <- 1/x1_2
root2_2 <- 1/x2_2

```

The Dynamic multiplier is **6.778241**. This is very high compared to the previous value, which explains that the shock value is pretty significant and this might be a point against the mean reversion.   
  
One of the characteristic roots has an absolute value greater than 1 (-0.5512492,1.451249). This means that this process is **not stationary**. From an intuition point of view, even though there was a small reversion in the AR(1) part ($phi_1$ = 0.9), the significantly positive AR(2) part ($phi_2$ = 0.8) doesn't allow the process to mean revert.  
  
  
#R Code
```{r eval=FALSE}
suppressMessages(library(xts))

stockreturns <- matrix(c(2012,-0.12,0.29,0.15,-0.03,2013,-0.03,0.15,0.03,0.05,2014,0.07,0.05,-0.1,0.25,2015,-0.06,-0.01,0.03,0.1,2016,0.33,-0.29,0.08,-0.18),nrow=5,byrow = TRUE)
stockreturns.xts <- as.xts(stockreturns[,-1],as.Date(paste0(stockreturns[,1],"-01-01")))
colnames(stockreturns.xts) <- c("Stock A","Stock B","Stock C","Stock D")

findMedian <- function(vec){
  vec.sorted <- sort(vec)
  if(length(vec.sorted) %% 2 == 0){
    (vec.sorted[length(vec.sorted)/2] + vec.sorted[(length(vec.sorted)/2)+1])/2
  }
  else{
    vec.sorted[length(vec.sorted)%/%2 + 1]
  }
}

##1a
#Store positions of winning stocks and convert it into xts
Winner.positions <- xts(t(apply(stockreturns.xts,1,function(x){which(x > findMedian(x))})),order.by=index(stockreturns.xts))

#PRINT WINNERS
Winner.names <- t(sapply(1:nrow(stockreturns.xts),
                        function(count){
                        #Print column name of calculated indices
                        colnames(stockreturns.xts[count,Winner.positions[count]])
                }))
Winner.names <- xts(Winner.names,index(stockreturns.xts))
Winner.names

#CALCULATE RETURNS
WinnerLoserReturnsFunc <- function(row){
                          #get previous winners using lag
                          previousWinners <- stats::lag(Winner.positions)[-1][index(row),]
                          #calculate mean of winners(containing previousWinners) and losers
                          c(mean(row[,previousWinners]),mean(row[,-previousWinners]))
                        }

WinnerLoserReturns <- t(sapply(1:nrow(stockreturns.xts[-1]),function(count)
                                    {
                                      WinnerLoserReturnsFunc(stockreturns.xts[-1][count,])
                                    }))

colnames(WinnerLoserReturns) <- c("Winner","Loser")
WinnerLoserReturns <- xts(WinnerLoserReturns,index(stockreturns.xts[-1]))
WinnerLoserReturns

#MEAN OF RETURNS
apply(WinnerLoserReturns,2,mean)

##1b

Market.weights <- c(0.3,0.2,0.353,0.147)
Market.returns <- stockreturns.xts%*%Market.weights

#MEAN MARKET 
mean(Market.returns)

LongShort.Returns <- WinnerLoserReturns[,'Winner'] - WinnerLoserReturns[,'Loser']
colnames(LongShort.Returns) <- "LongShort"
LongShort.Returns

#MEAN
LongShort.mean <- mean(LongShort.Returns)

#REGRESSION
reg <- summary(lm(LongShort.Returns ~ Market.returns[-1]))
betaEst <- reg$coefficients[2,1]

##2

Positions<- t(apply(stockreturns.xts,1,
                        function(timeVal){
                            sortedVal <- sort(coredata(timeVal))
                            sapply(sortedVal,function(elem){which(timeVal==elem)})
                          }))
Positions.xts <- xts(Positions,index(stockreturns.xts))
Positions.xts.lag <- stats::lag(Positions.xts)[-1]

QuartileSorts <- t(sapply(1:nrow(Positions.xts.lag),function(x){
  val <- stockreturns.xts[-1][x,Positions.xts.lag[x,]]
  colnames(val) <- NULL
  val
  }))

colnames(QuartileSorts) <- c("Sort1","Sort2","Sort3","Sort4")
QuartileSorts <- xts(QuartileSorts,index(Positions.xts.lag))
QuartileSorts

#MEAN
apply(QuartileSorts,2,mean)

#AR(p) Process  
##3a 
phi1 <- 1.1
phi2 <- -0.25

set.seed('1234')
ar.sim1 <-arima.sim(model=list(ar=c(phi1,phi2)),n=1000) 
acf(ar.sim1,lag.max = 20,main="Autocorrelation plot for AR(2)")

##3b
x1 = (phi1 + sqrt(phi1^2 + 4*phi2))/(-2 * phi2)
x2 = (phi1 - sqrt(phi1^2 + 4*phi2))/(-2 * phi2)

root1 <- 1/x1
root2 <- 1/x2

##3c
dynamicMult_1 <- phi1^6 + 5*phi1^4*phi2 + 6*phi1^2*phi2^2 + phi2^3

##3d
phi1 = 0.9
phi2 = 0.8

dynamicMult_2 <- phi1^6 + 5*phi1^4*phi2 + 6*phi1^2*phi2^2 + phi2^3

x1_2 = (phi1 + sqrt(phi1^2 + 4*phi2))/(-2 * phi2)
x2_2 = (phi1 - sqrt(phi1^2 + 4*phi2))/(-2 * phi2)

root1_2 <- 1/x1_2
root2_2 <- 1/x2_2
```
